\documentclass{article}

\usepackage{tikz}
\usepackage{parskip}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage{amsmath, empheq}
\usepackage[margin=3.75cm]{geometry}
\definecolor{mycolour}{RGB}{46,52,64}
\pagecolor{mycolour}
\color{white}

\title{\jobname}
\author{Eugenio Animali}

\begin{document}
\maketitle
\section{Life}
Full life of affairs, sports and enterprise. Suicide like his father.
Works for the International Red Cross during the first World War, on the border for Italy. In the ranks of the officers (lefttenant).
\section{A Farewell to Arms}
\subsection{Title}
Play on words, arms refers to:
\begin{itemize}
    \item
    weapons of war $\to$ war, death, disruption
    \item
    hugs, embraces $\to$ love, happiness, fulfilment
\end{itemize}
Union of these two opposite themes in the novel.

The beginning was inspired by his work in the Italian Alps. It was a time where most countries were monarchies, and there was a strong class divide. Tension between officers and simple soldiers. The situation was very bad in Italy because there was not much industrialisation and most soldiers were illiterate peasants, who did not speak italian; they only spoke their own dialects- leading to soldiers not following orders because they did not understand, and excecutions andministered by officers to punish them. The simple soldiers were abused constantly and brutaly. WWI was fought in the trenches and with charges; the higene, and conditions were horrible.
\subsection{Storyline}
Henry is an American officer, who had therefore a better relationship with the soldiers. As he goes to get some wine and cheese to bring to them, the enemy attacks and the building he is in collapses, and he is badly wounded. He is sent to a hospital in Milan (as Hemingway himself did) and falls in love with a nurse. When he is well again, he is sent back during the battle of Caporetto, great defeat for Italy. Some of the soldiers give themselves to the enemy because the motivation is all lost. Henry gives orders to continue and they do not obey, he must kill them. After this episode, he decides he does not want to do this anymore and escapes to Milan (Farewell to War) to reconcile with Cathrine, then with her to Lago di Como, but now he is exerted, and so if he is caught he will be excecuted. With the help of a friend they decide to cross Lake Como into Switzerland. In the mean time Cathrine discovers she is pregnant. Both Cathrine and the baby die. Final inner monologue. (Farewell to Love)
\subsection{Stile}
Short sentences, hardly any introspection, we only know what people think through what they say- all goes to place us in a realistic, warlike setting.

Rain falls on every important scene of difficulty $\to$ Symbol of Fate oppressing him.
\section{In the war}
Hemingway describes the situation he finds himself in as bombs fall. Full of onomatopoeia, short sentences. Hemingway finds Passini with his legs smashed. One was twitching
\end{document}