\documentclass{article}

\usepackage{tikz}
\usepackage{parskip}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage{amsmath, empheq}
\usepackage[margin=3.75cm]{geometry}
\definecolor{mycolour}{RGB}{46,52,64}
\pagecolor{mycolour}
\color{white}

\title{\jobname}
\author{Eugenio Animali}

\begin{document}
\maketitle

\section{themes}
Joyce uses all literary techniques except the omniscient narrator.

Although he studied in a jesuit college, Joyce was a free mind, and this is why he decided to escape Dublin for good.

Dublin - defined as ``the centre of paralysis'' dubliners are paralysed by the british empire because the irish are rebellious therefore oppressed. Their economy was heavily taxxed by the empire, and suffered hunger. The catholic church imposed its view of life and strict moral code in Ireland.

Every story is an individual failure which mirrors the universal failure which is the general human condition, particularly prominent in Dublin.

The characters do try to rebel and change their fate, but they are not strong, able or willing enough to make it happen.

Epiphany - the moment of realization of the characters. A sudden, unexpected understanding of the meaning of life. It always happens following an occurrence in the material world - union of realism and symbolism.

\section{Eveline}

Eveline is looking out of the window. We hear an inner monologue, the thoughts of Evenline, in third person. Memories invade her thoughts. Where now there are new houses, she used to play with other people's children, chased by her father with a stick. Now her father is more grumpy and her mother is dead. She will soon have to leave her house. She looks around her room, which she had never thought of having to leave.

Her fiancé is Frank, and they would go to live in Buenos Ayres, where he has a house. They used to meet after her work in the Stores, and go out to the opera. He sings and is fond of music. He is a sailor. He had started as a deck boy working for a pound a month.

She was initially happy to just have a boyfriend, but then she began to like her. Her father, however, does not appreciate their love because he had a quarrel with Frank once and from then the affair has continued in secret.

Now her father has become more gentle, and makes his children laugh.

Now she hears a song being played on a street organ outside. It is the same song as was played when her mother died, reminding her of the promise she made, to keep the family together.

Remembering her mother, she thinks about her craziness near the end of her life, which scares her.

When she finds Frank, she cannot recon with the idea of leaving and is paralysed, looking at Frank with a dead face among the crowd.
\section{The Dead}
Story of a traditional party held on the night of the epiphany on January 5th 1912.
\subsubsection{The Characters}
The Morkarn family

Miss Julie

Kate

Mary Jane, their niece, is a piano teacher

Gabriel and Gretta Comroy

Gabriel is the intellectual of the family thanks to his job in the university college, and is always the one to give the speech in family occations.

Gretta is not intellectual but affectionate. She represents Nora Joyce, James' sister.

The party is full of music and drinking, and --- doesn't want to sing, using the excuse of a sore throat, but he decides to seduce a girl at the end and she makes him sing.

Gabriel and Gretta go to a hotel, and Gabriel wants to make love but Gretta is not in the mood. She gets worked up, and she explains that when she would visit her grandmother in the summer in galway in her childhood, a boy called Michael Fury sang the song that was sung at the party. Gabriel gets jealous but she tells him Michael died. How did he die? "I think he died for me". He came to see her in the rain -i dont care if i die, i just want to see you-
\section{Ulysses}

\end{document}